---
title: "Averaging Data for Presenting"
output:
  html_document:
    df_print: paged
---

This script will go through the datasets for the different RCPs and take the different outputs, it will read them all in in order to make box plot and maps.

```{r Setup, echo=FALSE}
library(plotly)

# Set some variables for folder/file paths:
data_store <- "I:/SHETRAN_GB_2021/08_Analysis/Climatic-Drought/Outputs/SPEI_12/"
rcps <- c('01', '04', '05', '06', '07', '08', '09', '10', '11', '12', '13', '15')

# Set some more specifying the names of the outputs:
av_folder <- "Average Length of Drought - Gridded - "
av_file_name <- "drought_mean_length_raster_"

mx_folder <- "Maximum Length of Drought - Gridded - "
mx_file_name <- "drought_max_length_raster_"

pr_folder <- "Probability of Drought - Gridded - "
pr_file_name <- "drought_probability_raster_"

date_names = c("2000s", "2010s", "2020s", "2030s", "2040s", "2050s", "2060s")
file_dates = c("_19901201-20201130", "_20001201-20301130", "_20101201-20401130",
               "_20201201-20501130", "_20301201-20601130", "_20401201-20701130", 
               "_20501201-20801130")

aggregate_asscis = function(path){
  agg_asc = unlist(read.csv(read_path, skip=6, na.strings = -9999, header = FALSE, sep = " "))
  agg_asc = unname(agg_asc[!is.na(agg_asc)])
  agg_asc = agg_asc[!agg_asc==-9999]
  return(agg_asc)
}

```


# Drought Probability
```{r , echo=FALSE}
prob = list()
for (i in 1:7){
  x = c()
  for (rcp in rcps){
    read_path <- paste0(data_store, pr_folder, rcp, "/", pr_file_name, rcp, file_dates[i], ".asc")
    x = c(x, aggregate_asscis(read_path))
  }
  prob[[date_names[i]]]=x
}

```

```{r , echo=FALSE}
fig_pr <- plot_ly(type = "box")

for (i in date_names){
  fig_pr <- fig_pr %>% add_trace(x = i, y = prob[[i]], name = i)
}

fig_pr <- fig_pr %>% layout(title="Probability of a Month being a Drought", 
                      showlegend=TRUE, 
                      yaxis = list(title="Probability", range=c(0,1)))

fig_pr
```
 

# Mean Drought Length
```{r}
av = list()
for (i in 1:7){
  x = c()
  for (rcp in rcps){
    read_path <- paste0(data_store, av_folder, rcp, "/", av_file_name, rcp, file_dates[i], ".asc")
    x = c(x, aggregate_asscis(read_path))
  }
  av[[date_names[i]]]=x
}

```

```{r}
fig_av <- plot_ly(type = "box")

for (i in date_names){
  fig_av <- fig_av %>% add_trace(x = i, y = av[[i]], name = i)
}

fig_av <- fig_av %>% layout(title="Average Drought Length (months)", 
                      showlegend=TRUE, 
                      yaxis = list(title="Length (months)"))
fig_av
```


# Maximum Drought Length

```{r}
mx = list()
for (i in 1:7){
  x = c()
  for (rcp in rcps){
    read_path <- paste0(data_store, mx_folder, rcp, "/", mx_file_name, rcp, file_dates[i], ".asc")
    x = c(x, aggregate_asscis(read_path))
  }
  mx[[date_names[i]]]=x
}

```

```{r}
fig_mx <- plot_ly(type = "box")

for (i in date_names){
  fig_mx <- fig_mx %>% add_trace(x = i, y = mx[[i]], name = i)
}

fig_mx <- fig_mx %>% layout(title="Maximum Drought Length (months)", 
                      showlegend=TRUE, 
                      yaxis = list(title="Length (months)"))

fig_mx
```


# Maps of Statistics

Show the three statistics (probability, mean length, max length) for the UK for the baseline and 2050 periods.

Average these across the 12 RCPs or show all of them at once.

```{r}
# gridded_pr_2050
# Switch=TRUE
# for (rcp in rcps){
#   
#   read_path <- paste0(data_store, pr_folder, rcp, "/", pr_file_name, rcp, file_dates[6], ".asc")
#   
#   if(Switch==TRUE){
#     
#     Switch=FALSE
#     
#     h = read.csv(read_path, header = FALSE, sep = " ", nrows = 6)[1:2]
#     
#     x = read.csv(read_path, header = FALSE, sep = " ", skip = 6, na.strings = -9999)
#     
#     
#   }
#   
#   agg_asc = unlist(read.csv(read_path, skip=6, na.strings = -9999, header = FALSE, sep = " "))
#   agg_asc = unname(agg_asc[!is.na(agg_asc)])
#   agg_asc = agg_asc[!agg_asc==-9999]
#   
#   
#   x = c(x, aggregate_asscis(read_path))
# }
# mx[[date_names[i]]]=x
```


